---
title: "data_manipulation"
output: github_document
---

```{r}
library(tidyverse)
```

Import dataframes:
```{r}
litters_df = 
    read_csv("./data_import_examples/FAS_litters.csv", na = c("NA", ".", ""))

litters_df = 
    janitor::clean_names(litters_df)

pups_df = 
    read_csv("./data_import_examples/FAS_pups.csv", na = c("NA", "."), skip = 3)

pups_df = 
    janitor::clean_names(pups_df)
```

## 'select'

'select' selects variables from my dataframe.
```{r}
select(litters_df, group, litter_number, gd0_weight, pups_born_alive)
```

'select' select a specific range
```{r}
select(litters_df, group:gd_of_birth)
```

'select' select by removal
```{r}
select(litters_df, -pups_survive)
```

select to remove a ranges of variable together
```{r}
select(litters_df, -(group:gd_of_birth))
```

select by prefix (ie."gd"); also 'ends_with()', and 'contains()'
```{r}
select(litters_df, group, starts_with("gd"))
```

rearrange by selecting
```{r}
select(litters_df, starts_with("gd"), group, litter_number)
```

rearrange by selecting, keep everything as listed; 'relocate' works the same
```{r}
select(litters_df, litter_number, everything())

relocate(litters_df, litter_number)
```

rename by selecting
```{r}
rename(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

what if only select one variable...?
```{r}
select(litters_df, group)
```











