---
title: "Data Import"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

Let's import a dataset.

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv")
```

Look at your data!!

```{r}
names(litters_df)
```

Update the names in "litter_df".

```{r}
litters_df =
  janitor::clean_names(litters_df)
```

Look at the data!!

```{r}
litters_df
```

Sometimes skimming data is neat

```{r}
skimr::skim(litters_df)
```

## Fix the missingness

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv", na = c("NA", ".",""))
```


```{r}
pups_df = 
  read_csv("data_import_examples/FAS_pups.csv", 
           skip = 3, na = c("NA", ".","")
           )

pups_df = 
  janitor::clean_names(pups_df)
```

## Okay what about excel

CSVs are really great but soemtimes you get an excel file.
Remember to run Read_Excel library first.
```{r}
mlb_df = 
  read_excel("data_import_examples/mlb11.xlsx")
```

Import Lord of RIng word counts.

```{r}
fotr_df = 
  read_excel("data_import_examples/LotR_Words.xlsx", range = "B3:D6")
```

## Read SAS File
Remember to run Haven package first.
```{r}
pulse_df =
  read_sas("data_import_examples/public_pulse_data.sas7bdat")

pulse_df = 
  janitor::clean_names(pulse_df)
```

## Why hate read.csv so much?
It prints unorganize
```{r, eval = FALSE}
litters_df = 
  read.csv("data_import_examples/FAS_litters.csv")
```

## What about data exporting?

```{r}
write_csv(fotr_df,"data_import_examples/fotr_df.csv")
```






