---
title: "tidy_data"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)

options(tibble.print_min = 5)
```

This document will show how to tidy data.

## Pivot longer

shorten the columns and combine variables
` names_prefix` delete prefix among all variables
changing bl to 00m for consistency across visits and converting visit to a factor variable
```{r}
pulse_df = read_sas("./data_import_examples/public_pulse_data.sas7bdat")|>
  janitor::clean_names()|>
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi_score",
    names_prefix = "bdi_score_"
  )|>
  mutate(visit = replace (visit, visit == "bl", "00m")) |>
  relocate(id,visit)
```

Learning Assessment Code:
```{r}
litters_wide = 
  read_csv(
    "./data_import_examples/FAS_litters.csv",na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  pivot_longer(
    cols = gd0_weight:gd18_weight,
    names_to = "gd_time", 
    values_to = "weight") |>
  mutate(gd_time = case_match(gd_time, "gd0_weight" ~ 0, "gd18_weight" ~ 18))

```

## Pivot wider

Let's make up on analysis result table.
```{r}
analysis_result = 
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
  )

analysis_result
```

Pivot wider for human readability:
knitr::kable() produce a nicer table for reading in html format.
```{r}
pivot_wider(
  analysis_result, 
  names_from = "time", 
  values_from = "mean") |>
  knitr::kable()
```


## Bind tables.

Import/Read each table from excel:
```{r}
fellowship_ring = 
  readxl::read_excel("./data_import_examples/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data_import_examples/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data_import_examples/LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_king")
```


Stack tables up, using `bind_rows` function:
```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) |>
  janitor::clean_names() |>
  pivot_longer(
    female:male,
    names_to = "gender", 
    values_to = "words") |>
  mutate(race = str_to_lower(race)) |> 
  select(movie, everything()) 

lotr_tidy
```

## Joing FAS datasets

Import `litters` dataset.
```{r}
litter_df = 
  read_csv("./data_import_examples/FAS_litters.csv",na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_tx"), sep = 3) |>
  relocate(litter_number) |>
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    dose = str_to_lower(dose))
```


Import `pups` next:
```{r}
pups_df = 
  read_csv("./data_import_examples/FAS_pups.csv", na = c("NA", "."), skip = 3) |>
  janitor::clean_names() |>
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"),
    sex = as.factor(sex)) 
  
```

Join the datasets!

There are four major ways join dataframes x and y:
Inner: keeps data that appear in both x and y
Left: keeps data that appear in x
Right: keeps data that appear in y
Full: keeps data that appear in either x or y

Left joins are the most common, because they add data from a smaller table y into a larger table x without removing anything from x.
```{r}
fas_df = 
  left_join(pups_df, litter_df, by = "litter_number") |>
  select(litter_number, everything())

fas_df
```










